---
title: "compare"
author: "Gina"
format: 
  html: default
  gfm: default
editor: visual
---

```{r, echo = F, messages = F}
rm(list = ls())
suppressPackageStartupMessages(
  library(tidyverse)
)
source("00_conversions.R")

```

# Introduction to irrigation energy

There may be two sources of water:

1.  surface

2.  ground

Ground water will have a 'head' associated with the depth of the well. Both will have 'head' associated with the pump pressure.  

The main components defining the irrigation energy use are:

1.  The pump pressure and, if ground water, the well depth

2.  The energy source for the pumping/moving of water (diesel, electricity, solar, etc.)

3.  The amount of irrigation water applied

# Scenario

-   A flood irrigated field

-   Diesel powered pump

-   150 foot well

-   25 PSI pump

-   \$1 energy unit (to make back-calculating the amount of fuel easy)

-   Alfalfa

-   2.475 acre

-   68 ac-in applied per acre, 30 as ground, 38 as surface

-   The zipcode is used to auto-fill some NRCS values which we overwrite. It doesn't impact the calculations if you overwrite them. However I put in 93603.

    ```{r}
    well_depth_ft <- 150
    pump_press_psi <- 25
    water_applied_in_ac <- 30
    ```

# NRCS tool

The NRCS has a tool that will estimate the energy used for irrigation: <https://ipat.sc.egov.usda.gov/Default.aspx>

When using the above inputs, the estimated energy cost is \$54/ac. I take this to mean I will have used 54 gallons of diesel per acre:

```{r}
nrcs_gal_dies_ac <- 54

(nrcs_gal_dies_ha <- nrcs_gal_dies_ac * ac_per_ha)
```

According to Table 1 of FTM, 1 gallon of diesel contains 138,490 BTUs. We can calculate the number of BTUs we used per hectare, and change it to MJ/ha, multiply by 3 for the stand life:

```{r}
btu_per_gal_dies <- 138490

nrcs_mj_per_ha <- nrcs_gal_dies_ha * btu_per_gal_dies * mj_per_btu * 3

prettyNum(round(nrcs_mj_per_ha, 0), big.mark = ",", scientific = F)
```

Around 50,000 MJ per hectare for the ground water portion.

Just for ballpark estimates, Table 1 of FTM says there are 22.7 lbs of co2e per gallon of diesel (although this is from combustion of the fuel, only).

```{r}
lbsco2_per_gal_dies <- 22.7

nrcs_kgco2_ha <- nrcs_gal_dies_ac * ac_per_ha * lbsco2_per_gal_dies * kg_per_lb

prettyNum(nrcs_kgco2_ha, big.mark = ",", scientific = F)
```

Irrigation releases co2 on the order of Mg for the ground water portion.

# FTM equations

The total head is expressed as the sum of the head from the pump pressure and the pumping depth (Latex sucks in this environment):

head \[m\] = pumping pressure \[m\] + well depth \[m\]

To calculate the head we therefore just need to do some unit conversions.

```{r}
pump_press_m <- pump_press_psi * mhead_per_psi
well_depth_m <- well_depth_ft * m_per_ft

head_m <- pump_press_m + well_depth_m
print(round(head_m, 0))
```

This is then used in the pumping energy equation:

$$energy = \frac{H*C_{pump}*(acre-inches_{applied})*25.4 \frac{mm}{in} * (number-of-acres-in-field) * (0.404 \frac{ha}{acre})}{E_{pump}*E_{irrigation}*E_{gearhead}*E_{power-unit}}$$

Where:

$$H = head\\
C_{pump} = 0.0979\\
E_{pump} = 0.75\\
E_{irrigation} = 1\\
E_{gearhead} = 0.95\\
E_{power-unit} = 1$$

The units are very strange, and I cannot follow what is being canceled. Apparently this produces a value in MJs per hectare? The mm per inch conversion is very strange.

Constants provided by FTM (no citations):

```{r}
# constants, no idea of units
c_pump <- 0.0979
e_pump <- 0.75
e_irr <- 1
e_gear <- 0.95
e_power <- 1

```

Calculate, multiply by 3 for the stand life:

```{r}
(num1 <- head_m * c_pump * water_applied_in_ac * mm_per_in * ha_per_ac)
(den1 <- e_pump * e_irr * e_gear * e_power)
(ftm_mj_per_ha <- num1/den1 * 3)

```

Use this to compare to 01_energy calcs:

```{r}
res <- tibble(calc = c("nrcs ground", 
                "ftm ground"),
       btu_per_ha = c(nrcs_mj_per_ha,
                      ftm_mj_per_ha)) %>% 
  mutate_if(is.numeric, round, 0)
res
```
