---
title: "compare"
author: "Gina"
format: 
  html: default
  gfm: default
editor: visual
---

```{r, echo = F, messages = F}
rm(list = ls())
suppressPackageStartupMessages(
  library(tidyverse)
)
source("00_conversions.R")

```

# Introduction to irrigation energy

There may be two sources of water:

1.  surface

2.  ground

Ground water will have a 'head' associated with the depth of the well. Both will have 'head' associated with the pump pressure.  

The main components defining the irrigation energy use are:

1.  The pump pressure and, if ground water, the well depth

2.  The energy source for the pumping/moving of water (diesel, electricity, solar, etc.)

3.  The amount of irrigation water applied

The physics command a certain amount of energy. However, the energy source (fuel, electricity) is not converted 100% to work energy. The physics equations, and therefore FTM/UCANR equations, do not take this into account. However, the NRCS tool does. This exercise is to see how much that matters.

# Scenario 1

-   A flood irrigated field

-   Diesel powered pump

-   150 foot well

-   25 PSI pump

-   \$1 energy unit (to make back-calculating the amount of fuel easy)

-   Alfalfa

-   64 ac-in applied per acre each year

-   Get BTU/acre (which is what FTM reports results in)

-   According to Table 1 of FTM, 1 gallon of diesel contains 138,490 BTUs.

    ```{r}
    well_depth_ft <- 150
    pump_press_psi <- 25
    water_applied_in_ac <- 64
    ```

I used the FTM source for energy contained in a unit of diesel, and the BTUs per kwh is a conversion:

```{r}

btu_per_gal_dies <- 138490
btu_per_kwh

```

## NRCS tool

The NRCS has a tool that will estimate the energy used for irrigation: <https://ipat.sc.egov.usda.gov/Default.aspx>

```{r}
11.81 * btu_per_mj * l_per_gal
```

We can calculate the number of BTUs we used per acre:

```{r}

nrcs_gal_dies_ac <- 134
nrcs_btu_per_ac <- nrcs_gal_dies_ac * btu_per_gal_dies

prettyNum(round(nrcs_btu_per_ac, 0), big.mark = ",", scientific = F)
```

## FTM equations (these eqn units are f\*cked up)

Convert ac-in to mm-ha, and psi to kPa

```{r}
water_applied_mm_ha <- water_applied_in_ac * mm_per_in * ha_per_ac
pump_press_kpa <- pump_press_psi * kpa_per_psi
```

Use the constants assigned by the book and FTM:

```{r}
# constants
c_units <- 0.0979 #--no idea of units, book does not say
# working backwards, I think this is the gravity constant.
# But I don't get the same value they do. 
c_units2 <- 97.9 #--no idea of units, book does not say
#--some of these could change based on sprinkler/furrow etc. 
e_pump <- 0.75
e_irr <- 1
e_gear <- 0.95
e_power <- 1
```

Calculate the FTM values by hand:

NOTE: The book equation says it produces a value in 'MJ'. But there MUST be an area assumption built into it. They don't list the units of the constant.

```{r}
#--this is the eqn in the book
head_m <- 0.102 * pump_press_kpa + well_depth_ft * m_per_ft
num1 <- (c_units * water_applied_mm_ha * head_m)
den1 <- e_pump * e_irr * e_gear * e_power
manftm_mj_per_ha <- num1/den1
manftm_btu_per_ac <- manftm_mj_per_ha * btu_per_mj #* ha_per_ac
prettyNum(round(manftm_btu_per_ac, 0), big.mark = ",", scientific = F)
```

Fieldprint calculator value:

```{r}
ftm_btu_per_ac <- 5436395
```

## UCANR eqns

Calculate how many pounds of water you are lifting?

```{r}
gal_applied <- water_applied_in_ac * ft_per_in * gal_per_acft 
(lb_water_lifted <- gal_applied * lb_per_gal_water)
```

Multiply by how far you are lifting it in feet - this is foot pounds of energy. Change to BTUs

```{r}
(btu_used_to_lift <- lb_water_lifted * well_depth_ft * btu_per_ftlb)
```

1 psi of pressure is creating 2.31 feet of water head.

```{r}
(psi_depth_ft <- 25 * 2.31)
btu_used_to_pump <- lb_water_lifted * psi_depth_ft * btu_per_ftlb
```

Total energy used on this acre:

```{r}
(ucanr_btu_per_ac <- btu_used_to_lift + btu_used_to_pump)
```

# Scenario 2

-   Electric pump

-   \$0.10 per kwh

## NRCS tool

```{r}
nrcs_cost_elec <- 190
nrcs_kwh_ac <- nrcs_cost_elec / 0.10
nrcs_btu_per_ac2 <- nrcs_kwh_ac * btu_per_kwh

prettyNum(round(nrcs_btu_per_ac2, 0), big.mark = ",", scientific = F)
```

## FTM equations (these eqn units are f\*cked up)

The equations in FTM don't change based on the energy source, so calculating it by hand wouldn't change anything. This is confirmed in that the value doesn't change when I change the source.

Fieldprint calculator value:

```{r}
ftm_btu_per_ac2 <- 5436395
```

## UCANR eqns

These calculations won't change:

```{r}
(ucanr_btu_per_ac2 <- ucanr_btu_per_ac)
```

# 

# Summary

Use this to compare to 01_energy calcs:

```{r}
res <- tibble(calc = c(
                "nrcs 64 ground", 
                "ftm-manual 64 ground",
                "ftm 64 ground",
                "ucanrs 64 ground",
                "ucanrs 64 ground low eff"),
       btu_per_ac_diesel = c(nrcs_btu_per_ac,
                      manftm_btu_per_ac,
                      ftm_btu_per_ac,
                      ucanr_btu_per_ac,
                      ucanr_btu_per_ac/(0.75*0.95)),
       btu_per_ac_electricity = c(nrcs_btu_per_ac2,
                      manftm_btu_per_ac,
                      ftm_btu_per_ac,
                      ucanr_btu_per_ac,
                      ucanr_btu_per_ac/(0.75*0.95))) %>% 
  mutate_if(is.numeric, round, 0)
res
```

Simplify the results to write.

```{r}
res2 <- 
  res %>% 
  mutate(methodology = case_when(
    grepl("nrcs", calc) ~ "NRCS Energy Tool",
    grepl("ftm", calc) ~ "Field to Market Online Tool",
    grepl("ucanrs", calc) ~ "Univ Cal ANRS Equations"
  )) %>% 
  filter(calc %in% c("nrcs 64 ground", "ftm 64 ground", "ucanrs 64 ground low eff")) %>% 
  select(methodology, btu_per_ac_diesel, btu_per_ac_electricity)

res2 %>% write_csv("compare-irr-energy-methods-summary.csv")
```
