---
title: "compare"
author: "Gina"
format: 
  html: default
  gfm: default
editor: visual
---

```{r, echo = F, messages = F}
rm(list = ls())
suppressPackageStartupMessages(
  library(tidyverse)
)
source("00_conversions.R")

```

# Introduction to irrigation energy

There may be two sources of water:

1.  surface

2.  ground

Ground water will have a 'head' associated with the depth of the well. Both will have 'head' associated with the pump pressure.  

The main components defining the irrigation energy use are:

1.  The pump pressure and, if ground water, the well depth

2.  The energy source for the pumping/moving of water (diesel, electricity, solar, etc.)

3.  The amount of irrigation water applied

# Scenario

-   A flood irrigated field

-   Diesel powered pump

-   150 foot well

-   25 PSI pump

-   \$1 energy unit (to make back-calculating the amount of fuel easy)

-   Alfalfa

-   64 ac-in applied per acre each year

-   3 year stand

-   Get BTU/acre (which is what FTM reports results in)

-   According to Table 1 of FTM, 1 gallon of diesel contains 138,490 BTUs.

    ```{r}
    well_depth_ft <- 150
    pump_press_psi <- 25
    water_applied_in_ac <- 64
    btu_per_gal_dies <- 138490
    ```

# NRCS tool

The NRCS has a tool that will estimate the energy used for irrigation: <https://ipat.sc.egov.usda.gov/Default.aspx>

Saved output in We can calculate the number of BTUs we used per acre:

```{r}

nrcs_gal_dies_ac <- 134
nrcs_btu_per_ac <- nrcs_gal_dies_ac * btu_per_gal_dies

prettyNum(round(nrcs_btu_per_ac, 0), big.mark = ",", scientific = F)
```

# FTM equations (these eqn units are f\*cked up)

Convert ac-in to mm-ha, and psi to kPa

```{r}
water_applied_mm_ha <- water_applied_in_ac * mm_per_in * ha_per_ac
pump_press_kpa <- pump_press_psi * kpa_per_psi
```

Use the constants assigned by the book and FTM:

```{r}
# constants
c_units <- 0.0979 #--no idea of units, book does not say
# working backwards, I think this is the gravity constant.
# But I don't get the same value they do. 
c_units2 <- 97.9 #--no idea of units, book does not say
#--some of these could change based on sprinkler/furrow etc. 
e_pump <- 0.75
e_irr <- 1
e_gear <- 0.95
e_power <- 1
```

Calculate the FTM values by hand:

NOTE: The book equation says it produces a value in 'MJ'. But there MUST be an area assumption built into it. They don't list the units of the constant.

```{r}
#--this is the eqn in the book
head_m <- 0.102 * pump_press_kpa + well_depth_ft * m_per_ft
num1 <- (c_units * water_applied_mm_ha * head_m)
den1 <- e_pump * e_irr * e_gear * e_power
manftm_mj_per_ha <- num1/den1
manftm_btu_per_ac <- manftm_mj_per_ha * btu_per_mj #* ha_per_ac
prettyNum(round(manftm_btu_per_ac, 0), big.mark = ",", scientific = F)
```

Fieldprint calculator value:

```{r}
ftm_btu_per_ac <- 5436395
```

# UCANR eqns

Calculate how many pounds of water you are lifting?

```{r}
gal_applied <- water_applied_in_ac * ft_per_in * gal_per_acft 
(lb_water_lifted <- gal_applied * lb_per_gal_water)
```

Multiply by how far you are lifting it in feet - this is foot pounds of energy. Change to BTUs

```{r}
(btu_used_to_lift <- lb_water_lifted * well_depth_ft * btu_per_ftlb)
```

1 psi of pressure is creating 2.31 feet of water head.

```{r}
(psi_depth_ft <- 25 * 2.31)
btu_used_to_pump <- lb_water_lifted * psi_depth_ft * btu_per_ftlb
```

Total energy used on this acre:

```{r}
(ucanr_btu_per_ac <- btu_used_to_lift + btu_used_to_pump)
```

# Summary

Use this to compare to 01_energy calcs:

```{r}
res <- tibble(calc = c(
                "nrcs 64 ground", 
                "ftm-manual 64 ground",
                "ftm 64 ground",
                "ucanrs 64 ground",
                "ucanrs 64 ground low eff"),
       btu_per_ac = c(nrcs_btu_per_ac,
                      manftm_btu_per_ac,
                      ftm_btu_per_ac,
                      ucanr_btu_per_ac,
                      ucanr_btu_per_ac/(0.75*0.95))) %>% 
  mutate_if(is.numeric, round, 0)
res
```

# 
